Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary

### 2017-04-12 inst/documents/paper2/paper2_resistance_mechanisms_mixtures.Rmd
paper writing

---
title: "paper2_resistance_mechanisms_mixtures"
output: pdf_document
#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
bibliography: library.bib
csl: biomed-central.csl
#header-includes:
#    - \usepackage{setspace}
#    - \doublespacing
#    - \usepackage{lineno}
#    - \linenumbers
---

@@

### version 11

#Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary.

@@

Target : Malaria journal

THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the express permission of one of the authors. Comments and suggestions welcome

@@

#### Implications
_todo get more of [@Roush1989] in here_

@@

_low exposure could be achieved by having active ingredients with some repellency._

_todo get Birget and Koella in & link to exposure_
'whereas repellency is beneficial to delay the evolution of resistance, other models have shown that it decreases the population-level protection of the insecticide.'

_todo, where to put this, earlier ?_
This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance. Other factors also need to be considered. On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease. On the negative side a mixture is more expensive.

@@

We have developed flexible, modern model that can replicate previous modelling work and outputs that help us understand the mechanisms by which different insecticide application strategies may be favoured. For assessing the relative benefits of mixtures and sequences we have demonstrated the likely importance of insecticide effectiveness and exposure and costs of resistance. High insecticide effectiveness favours mixtures through the protective effect of one insecticide on the other. High exposure favours sequences through reducing the otherwise high selection pressures on both resistances together. High costs of resistance can favour mixtures because the selection pressures on both resistances are reduced. Similar processes will occur in other strategies such as rotations and mosaics and we are starting to investigate them. Predictions from the model can be tested against data from trials of new insectides to help advance guidelines on good practice.

@@

##### Ethics approval and consent to participate
Not applicable
##### Consent for publication
Not applicable
##### Availability of data and material

>>>>

---
title: "Insecticide resistance evolution with mixtures and sequences : a model-based graphical introduction and summary"
output: 
  pdf_document:
    #citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex

#output: word_document
author: "Andy South and Ian Hastings"
date: "`r Sys.Date()`"
bibliography: library.bib
csl: biomed-central.csl
fontsize: 11pt
spacing: double
# below worked in test_linenumbers.Rmd
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
---

@@

### version 12

@@

Target : Malaria journal (other suggestions welcome)

THIS IS A CONFIDENTIAL COPY OF THE MS. Please do not circulate without the permission of one of the authors. Comments and suggestions welcome

@@

#### Implications

@@

_TODO bits to incorporate into discussion :_

_todo more reference to [@Roush1989] who made many of these points 25 years ago_

_Insecticides with some repellency will reduce exposure to the insecticide and thus, as shown here, are likely to slow the evolution of resistance and favour mictures over sequences. This would need to be balanced with a likely loss in protection to the population as a whole provided by the insecticide [@Birget2015]._

_This model can explore when a mixture strategy may be favoured over a sequence purely for the evolution of resistance. Other factors also need to be considered. On the positive side a mixture is likely to kill more vectors and may also limit their ability to transmit the disease. On the negative side a mixture is more expensive._

@@

We have developed flexible, modern model that can replicate previous modelling work and outputs that help us understand the mechanisms by which different insecticide application strategies may be favoured. For assessing the relative benefits of mixtures and sequences we have demonstrated the likely importance of insecticide effectiveness and exposure and costs of resistance. High insecticide effectiveness favours mixtures through the protective effect of one insecticide on the other. High exposure favours sequences because of the higher selection pressures on both resistances together in a mixture. High costs of resistance can favour mixtures because the selection pressures on both resistances are reduced. Similar processes will occur in other strategies such as rotations and mosaics and we are starting to investigate them. Predictions from the model can be tested against data from trials of new insectides to help advance guidelines on good practice.

@@

##### Ethics approval and consent to participate
Not applicable

##### Consent for publication
Not applicable

##### Availability of data and material
